Breadth First Searches Over Evolving Graphs

### 2016-01-27 paper.tex
Rewrite block adjacency matrix section

The proof of Theorem~\ref{thm:correct} provided a construction for representing
an evolving graph $G_n$ by a static graph $G$ with nodes corresponding to active
nodes of $G_n$. The adjacency matrix $\mathbf A$ of $G$ then has block structure
which is useful for understanding the nature of the $\odot$ operation.

@@

\begin{subequations}
\begin{align}
  \label{eq:5}
 \big\langle (A^{[1]})^Tc_1,   (A^{[2]})^T\odot c_1, \ldots,  (A^{[n]})^T\odot c_1  \big\rangle  \\
\big\langle (A^{[2]})^Tc_{2}, \ldots, (A^{[n]})^T\odot c_{2} \big\rangle \\
 \ldots  \\
  \big\langle (A^{[n]})^Tc_n \big\rangle
\label{eq:6}
\end{align}
\end{subequations}
%
where $c_i = (A^{[i]})^Te_k$. Summing resultant vectors that share the same
time stamp, we obtain vectors whose nonzero elements have indexes labeled by
the forward neighbors of the nodes computed at step $1$.

Define the block triangular adjacency matrix
\[
\bm{A_n}^T =
\begin{bmatrix}
(A^{[1]})^T &   0            & \cdots  &  & 0 \\
\sigma(A^{[2]})^T & (A^{[2]})^T & 0         &\cdots  & \vdots \\
\sigma(A^{[3]})^T & \sigma(A^{[3]})^T & (A^{[3]})^T & \cdots \\
   \vdots              &                              &      \vdots            &  &  \vdots \\
\sigma(A^{[n]})^T &                  &       \cdots                      &  & (A^{[n]})^T
\end{bmatrix},
\]
where $\sigma(A^{[i]})^Tb = (A^{[i]})^T\odot b$, and the block vector
\[
\bm{b}^T = [b^T, 0, \cdots, 0  ].
\]
Then, \eqref{eq:out_ne} is equivalent to the matrix--vector product
$\bm{A_n}^T\bm{b}$ and the summation over \eqref{eq:5}-\eqref{eq:6} is equivalent
to $\bm{A_n}^T(\bm{A_n}^T\bm{b})$.

>>>>

The proof of Theorem~\ref{thm:correct} provides a construction for representing
an evolving graph $G_n$ by a static graph $G$ with nodes corresponding to active
nodes of $G_n$. It turns out that the block structure of $G$ is useful for
understanding the nature of the $\odot$ operation.

@@

\begin{subequations}
\begin{align}
  \label{eq:5}
 \big\langle (A^{[1]})^Tc_1,   (A^{[2]})^T\odot c_1, \ldots,  (A^{[n]})^T\odot c_1  \big\rangle  \\
\big\langle (A^{[2]})^Tc_{2}, \ldots, (A^{[n]})^T\odot c_{2} \big\rangle \\
 \ldots  \\
  \big\langle (A^{[n]})^Tc_n \big\rangle
\label{eq:6}
\end{align}
\end{subequations}
%
where $c_1 = (A^{[1]})^Te_k$ and $c_i = (A^{[i]})^T \odot e_k$ for $i>1$.
Summing resultant vectors that share the same
time stamp, we obtain vectors whose nonzero elements have indexes labeled by
the forward neighbors of the nodes computed at step $1$.

Compare this with the matrix

\[
\bm M_n = \begin{bmatrix}
A^{[t_1]} & M^{[t_1, t_2]} & ... & M^{[t_1, t_n]} \\
        0 & A^{[t_2]}      & ... & M^{[t_2, t_n]} \\
                           &   ... \\
        0 & 0 & ... & A^{[t_n]}
\end{bmatrix}
\]
%
where $M^{[t_i, t_j]}$ is the matrix whose rows are labeled by $V^{[t_i]}$ and
columns are labeled by $V^{[t_j]}$, and whose entries are
\[
M^{[t_i, t_j]}_{uv} =
\begin{cases}
1 & \mbox{if $(u,v)\in E^\prime$,} \\
0 & \mbox{otherwise.}
\end{cases}
\]
%
The adjacency matrix blocks $A^{[t]}$ encode the edge set $\tilde E^{[t]}$,
whereas the off-diagonal blocks $M^{[t_i, t_j]}$ together encode the edge set
$E^\prime$ which capture temporal paths with subsequences of the form
$\langle (v, t_i), (v, t_j) \rangle$. Then $\bm M_n$ is the adjacency matrix of
the graph $(\cup_t V^{[t]}, E)$, which is the graph $G$ together with all the
inactive nodes. From the definition, $\bm M_n$ has nonzero
entries only in rows and columns that correspond to active nodes $V$, and so
retaining only these rows and columns corresponding to $V$ produces the
adjacency matrix $\bm A_n$ of $G = (V, E)$.

The off-diagonal blocks $M^{[t_i, t_j]}$ provide an explicit matrix representation
for the $\odot$ product in that $(M^{[t_i, t_j]})^T b = A^T \odot b$. The upper
triangular structure of $\bm M_n$ (and hence $\bm A_n$) reflects the causal
nature of temporal paths in that they cannot go backward in time.

The BFS algorithm presented above can therefore be interpreted as
computing the sequence of matrix--vector products $\bm b$, $\bm A^T \bm b$,
$\bm (A^T)^2 \bm b$, formed by applying successive monomials of $\bm A^T$ to the
block vector
\[
\bm b^T = [b^T, 0, \cdots, 0 ].
\]

For the example of Figure~\ref{fig:eg_shortest_path}, we have
\begin{align*}
V & = \{(1,\!t_1), (2,\!t_1), (1,\!t_2), (3,\!t_2), (2,\!t_3), (3,\!t_3)\},\\
\cup_t\!E^{[t]} &= \{((1,\!t_1), (2,\!t_1)), ((1,\!t_2), (3,\!t_2)), ((2,\!t_3), (3,\!t_3))\},\\
E'  &= \{((1,\!t_1), (1,\!t_2)), ((2,\!t_2), (2,\!t_3)), ((3,\!t_2), (3,\!t_3))\}.
\end{align*}
In the order specified for $V$, the adjacency matrix of $G$ is then
\[
\bm A_3 = \begin{bmatrix}
0 & 1 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 1 \\
0 & 0 & 0 & 0 & 0 & 0
\end{bmatrix}
\]

Starting from the vector $\bm b = e_1$, the sequence of iterates is then
\[
\left\langle
\begin{bmatrix}1 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix},
\begin{bmatrix}0 \\ 1 \\ 1 \\ 0 \\ 0 \\ 0 \end{bmatrix},
\begin{bmatrix}0 \\ 0 \\ 0 \\ 1 \\ 1 \\ 0 \end{bmatrix},
\begin{bmatrix}0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 2 \end{bmatrix},
\begin{bmatrix}0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix},
\dots
\right\rangle.
\]

We see that $\bm (A^T)^2$ correctly encodes the collection of products in
\eqref{eq:5}-\eqref{eq:6}. Furthermore, $\bm (A^T)^3 \bm b$ correctly counts the
two allowed temporal paths from $(1, t_1)$ to $(3, t_3)$, and that the
off-diagonal structure encoded in $E'$ and the $M^{[t,t']}$ blocks are critical
to obtaining the correct count.
